---
title: "outbreaks"
output: html_document
date: '2022-03-23'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(pak)
library(readxl)
```

## Pre-course: EDA

```{r eda, options}
read_excel("data/NationalOutbreakPublicDataTool.xlsx", sheet = 1) %>%
  clean_names() 

o <- read_excel("data/NationalOutbreakPublicDataTool.xlsx", sheet = 1) %>%
  janitor::clean_names()
```

Initial questions as a project author:

* how many etiologies are there?
* how many "primary mode" types are there?
* which state has the most infections?
* which outbreak was the most lethal? (i.e. largest deaths/infections ratio)

### Other thoughts

* What if week 2 was "acquaint yourself with the data" week. It would cover:
  * importing
  * ggplot basics
  * janitor functions
  * encourage students to check out EDA packages like `DataExplorer`
  
  This alleviates having to cram both "Import" and all of ggplot into a single week.
  
* one constraint we have traditionally imposed on ourselves is that each week's lessons must be thematically related. It might make more sense if, in a single week, you are taught a little bit of ggplot and a little bit of dplyr. That way you can use both to create something interesting. The next week you do a bit more of those same two topics. 

* But one problem with the above is that it makes it hard for students to know how far to go with an extension after the first week: How do they know that their extension content be taught in a lesson the next week. 
    + Is it a bad thing ?
    + Is it *better* if this didn't happen?

## Week 1: R Basics

* Objects & Functions
* Vectors
* Strings and Arguments
* Tables
* (Logicals? from Noodle)

```{r}
unique(o$etiology) %>% length() # 679 of these
hist(o$illnesses) # heeeeeavily right skewed, would need to log transform
hist(o$year) # left-skewed. Most cases occur after 1995.
```

## Week 2: Import and Visualize

* Packages
* Importing data
* Import excel files
* ggplot2 intro
* Aesthetics
* Optional: Facets

**Milestone:**
* Read in the data as excel spreadsheet
* Recast week 1 plot with ggplot2.
* Any other file to read in (CSVs) available for students to explore for their extension?
* Would love if students could explore the janitor package and use `clean_names()`

```{r}
data <- read_excel("data/NationalOutbreakPublicDataTool.xlsx", sheet = 1) %>%
  janitor::clean_names()

data
```


## Week 3: Transform tables

* Logicals?
* 
